---
title: "Result - disaster"
author: "Yihui Xie"
date: "5/2/2022"
output: html_document
---

# result - disaster

```{r,echo=FALSE, message=FALSE, results='hide'}
library(ggplot2)
library(tidyverse)
library(ggrepel)
library(cowplot)
library(GGally)
library(parcoords)

disaster <- read.csv("/Users/zshong/Desktop/Disaster_analysis/data/disaster.csv")
```

## Analysis of the relationship between death and economic loss

```{r,warning=FALSE,warning=FALSE, message=FALSE,fig.width=10}
plot.data1 <- disaster %>%
  select(Country,Total.Deaths,Total.Damages.Adjusted,Continent) %>%
  group_by(Country) %>%
  summarize(Deaths = sum(Total.Deaths),Adjusted.Damages = sum(Total.Damages.Adjusted),
            disaster.Num =length(Total.Deaths),Continent = unique(Continent))

# plot 1
p1.1 <- plot.data1 %>%
  ggplot( aes(Deaths, Adjusted.Damages, size = disaster.Num, color=Continent)) +
  geom_point() +
  theme(legend.position = "none") +
  geom_text_repel(aes(y=Adjusted.Damages,x=Deaths,label=Country))

# plot 2
p1.2 <-ggplot(plot.data1,aes(log(Deaths), log(Adjusted.Damages), size = disaster.Num, color=Continent, text = Country)) +
  geom_point() +
  theme(legend.position="none")+
  geom_abline(intercept = 5)

# Set two plots on one canvas and share one legend.
prow <- plot_grid(p1.1,p1.2,labels = c("plot 1","plot 2"),label_size = 12,
  label_x = 0, label_y = 0,
  hjust = -0.5, vjust = -0.5)

legend_bottom <- get_legend(
  p1.1 + 
    guides(color = guide_legend(nrow = 2)) +
    theme(legend.position = "bottom")
)

plot_grid(prow, legend_bottom, ncol = 1, rel_heights = c(1, .15))

```

We calculated the cumulative economic losses, deaths, and occurrences of disasters for each country from 1900 to 2022, and plotted the economic losses and deaths as y and x, respectively, as a scatterplot "figure 1", where the size of the dots correlates with the occurrence of disasters , and the colors are divided according to the continent (this data unifies South and North America into the Americas). Through figure 1, we can see that most of the countries are clustered in a small range, and only a few countries are very significant as outliers, such as the United States, Japan, China, India, Russia. However, these countries did not show a direct proportional trend in the number of deaths and economic losses. Among them, although the United States and Japan have fewer deaths, the economic losses are very large. While China, India and Russia suffered less economic losses, they have a  large number of deaths. We believe that the reason for this phenomenon is that countries with developed economies will devote more resources to personnel protection, so even if the United States has a similar number of disasters as China, the death rate is far less than that of China. In contrast, the economic losses caused by disasters in these economically developed countries will be much greater than those in other countries. And countries like China, Russia and India are the most populous countries in the world, so that's why they have the highest number of disaster deaths.

Then we log the deaths and economic losses, then make the scatterplot "figure 2". From this graph, we can clearly see that each continent forms a certain cluster, and all countries have more economic loss ratios than death ratios (All countries are above line y=x). Countries in Africa have higher rates of death than economic losses compared to other continents. The second is Asian countries, which are generally distributed online, but the outlier countries, China, India and Bangladesh, are offline due to their large population. Japan is above the line because of its developed economy and small population. Above Asia are the countries of the Americas, Europe and Oceania. Among them, the United States' economic loss rate far exceeds the death rate due to its developed economy, and Russia has become the only European country below the line because of its large population. Through this graph, we believe that the more developed the economy is, the greater the proportion of economic losses than deaths. In most developing countries in Africa, for example, the ratio of deaths to economic losses is higher than in the Americas and Europe. From this we can see that developed countries attach more importance to people's protection.

## Analysis of the relationship between number of disaster, Deaths and Damages

In this section, we want to study the relationship between death and property damage and the number of disasters. We group and integrate the data according to the Disaster.Type in the data. Then do barplot with Disaster.Type and sort from large to small for the number of disasters, the number of deaths and property damage.

```{r}
plot.data2 <- disaster %>%
  select(Disaster.Type,Total.Deaths,Total.Damages.Adjusted,Disaster.Group)%>%
  group_by(Disaster.Group,Disaster.Type) %>%
  summarize(Death = sum(Total.Deaths),Damages = sum(Total.Damages.Adjusted),Num.disaster = length(Total.Deaths)) %>%
  ungroup()

plot.data2 %>%
  mutate(Disaster.Type = fct_reorder(Disaster.Type, Num.disaster)) %>%
  ggplot(aes(x=Disaster.Type, y=Num.disaster)) +
    geom_bar(stat="identity", width=.4) +
    coord_flip() +
    xlab("") +
    theme_bw()
```

The above plot is a bar plot of the frequency of disasters. We can see that there are three orders of magnitude for the number of disasters. The first magnitude exceeded the other disasters by more than three times, and they were "Transport accident", "Flood" and "Storm". There are four kinds of disasters in the second magnitude, the values between them are very similar and more than twice the number of remaining disasters, they are earthquake, industrial accident, miscellaneous accident and epidemic. There are 12 catastrophes in the last magnitude, all of which occur relatively infrequently.

```{r}
plot.data2 %>%
  mutate(Disaster.Type = fct_reorder(Disaster.Type, Death)) %>%
  ggplot( aes(x=Disaster.Type, y=Death)) +
    geom_bar(stat="identity", width=.4) +
    coord_flip() +
    xlab("") +
    theme_bw()
```

The figure above is a bar chart of the number of deaths caused by a disaster. We can see that there are many disasters with very few or even zero deaths, so we only observed six disasters with significant numbers. They are Drought, Epidemic, Flood, Complex Disaster, Earthquake and Storm. Comparing the above disaster frequency chart, we will find that the above five disasters do not completely belong to the first two orders of magnitude, and the frequency of complex disasters and droughts is actually much lower than that of the other three disasters. So we can think that deaths are not completely positively correlated with the frequency of disasters

```{r}
plot.data2 %>%
  mutate(Disaster.Type = fct_reorder(Disaster.Type, Damages)) %>%
  ggplot(aes(x=Disaster.Type, y=Damages)) +
    geom_bar(stat="identity", width=.4) +
    coord_flip() +
    xlab("") +
    theme_bw()
```

The figure above is a bar chart of economic losses caused by a disaster. We can see that there are many disasters with little or no economic loss, so we only observed three very significant disasters and four relatively significant disasters. Very notable disasters are storms, floods and earthquakes. Relatively notable disasters are droughts, wildfires, extreme temperatures, and industrial accidents. Comparing to the above disaster death map, we will find that except for floods and earthquakes, disasters that cause a large number of deaths are not necessarily accompanied by high economic losses. Comparing the above disaster frequency map, we will find that floods, storms, earthquakes and industrial accidents are indeed frequent disasters, and most of these disasters that cause high economic losses are natural disasters.

```{r}
cols <- character(nrow(plot.data2))
cols[] <- "black"

cols[plot.data2$Disaster.Group == "Natural"] <- "blue"
cols[plot.data2$Disaster.Group == "Technological"] <- "red"
cols[plot.data2$Disaster.Group == "Complex Disasters"] <- "orange"
```


```{r}
plot.data2.1 <- plot.data2 %>%
  select(Death,Damages,Num.disaster) %>%
  mutate(Death = log10(Death),Damages = log10(Damages),Num.disaster = log10(Num.disaster))

pairs(plot.data2.1,col=cols)
```

In order to further verify the relationship between disaster frequency and death and economic loss, we have drawn a pairwise plot, in which blue points represent natural disasters, red points represent technological disasters, and orange points represent complex disasters. . Through this figure, we can find that there is a certain correlation trend among the frequency of disasters, deaths and economic losses.

```{r}
target.Country = c('China','US','Russia','India','Brazil','Japan','France','Italy','South Africa','Egypt')

plot.data3 <- disaster %>%
  filter(Country %in% target.Country) %>%
  filter(Disaster.Group %in% c("Natural","Technological")) %>%
  group_by(Country,Disaster.Group) %>%
  summarize(Death = log(sum(Total.Deaths)),Damages = log(sum(Total.Damages.Adjusted)),Num.disaster = log(length(Total.Deaths)))

parcoords(plot.data3,
          color = list(
            colorBy = "Country",
            colorScale = "scaleOrdinal",
            colorScheme = "schemeCategory10"),
          rownames = F,
          brushMode = "1D-axes",
          reorderable = T,
          queue = T,
          withD3 = T)

```

To further verify our conclusions, we selected ten typical countries for further analysis, they are "China', 'US', 'Russia', 'India', 'Brazil', 'Japan', 'France', ' Italy', 'South Africa' and 'Egypt'. Among them, we choose "China', 'US', 'Russia', 'India', 'Japan', 'France', 'Italy' because these countries are relatively significant points in the previous scatter plot analysis. The choice of 'South Africa' and 'Egypt' is because South America and Africa are vast and have large populations, so we selected the countries that are representative of the continent for analysis.

We calculated deaths, economic losses and disaster frequencies for these ten typical countries according to different types of disasters and used these data to make interactive parallel coordinates plots. Through the above figure, we can find that there are certain clusters of deaths caused by technological disasters , and the deaths caused by natural disasters are very scattered. Moreover, the economic losses caused by technological disasters are often smaller than those caused by natural disasters. We think this may be related to the fact that technological disasters are generally smaller in scale than natural disasters.